---
title: "Primary trial outcome"
author: Laura Symul, Laura Vermeren, Susan Holmes
format: 
   html:
     page-layout: full
     code-fold: true
     toc: true
     toc-location: left
     toc-depth: 5
     embed-resources: true
execute:
  cache: true
  warning: false
knitr:
  opts_chunk:
    out.width: "100%"
editor: source
---



### Colonization by week 5

```{r}

col_by_week5 <- 
  mae@colData |> as_tibble() |> 
  select(pid, site, randomized, arm, ITT, mITT, PP) |> 
  distinct() |> 
  filter(randomized, mITT) |> 
  left_join(
    mae[["col_LBP_mg"]] |> 
      as_tibble() |> 
      dplyr::left_join(
        mae@colData |> as_tibble() |> select(uid, pid, visit_code),
        by = join_by(.sample == uid)
      ) |> 
      dplyr::filter(visit_code == "1500", .feature == "colonized_LBP_by_mg")
  ) |> 
  mutate(LBP_colonization_by_week5 = outcome |> replace_na(FALSE)) 
  

# by qPCR 

col_by_week5_qPCR <-
  mae@colData |> as_tibble() |> 
  select(pid, site, randomized, arm, ITT, mITT, PP) |> 
  distinct() |> 
  filter(randomized, mITT) |> 
  left_join(
    mae[["col_LBP_qPCR"]] |> 
      as_tibble() |> 
      dplyr::left_join(
        mae@colData |> as_tibble() |> select(uid, pid, visit_code),
        by = join_by(.sample == uid)
      ) |> 
      dplyr::filter(visit_code == "1500", .feature == "colonized_LBP_by_qpcr")
  ) |> 
  mutate(LBP_colonization_by_week5 = outcome |> replace_na(FALSE)) 


detected_by_week5_qPCR <-
  mae@colData |> as_tibble() |> 
  select(pid, site, randomized, arm, ITT, mITT, PP) |> 
  distinct() |> 
  filter(randomized, mITT) |> 
  left_join(
    mae[["col_LBP_qPCR"]] |> 
      as_tibble() |> 
      dplyr::left_join(
        mae@colData |> as_tibble() |> select(uid, pid, visit_code),
        by = join_by(.sample == uid)
      ) |> 
      dplyr::filter(visit_code == "1500", .feature == "LBP_detected_by_qpcr")
  ) |> 
  mutate(LBP_detected_by_week5 = outcome |> replace_na(FALSE)) 


# crispatus dominance 

crisp_dom_by_5week <- 
  mae@colData |> as_tibble() |> 
  select(pid, site, randomized, arm, ITT, mITT, PP) |> 
  distinct() |> 
  filter(randomized, mITT) |> 
  left_join(
    mae[["col_crispatus_mg"]] |> 
      as_tibble() |> 
      dplyr::left_join(
        mae@colData |> as_tibble() |> select(uid, pid, visit_code),
        by = join_by(.sample == uid)
      ) |> 
      dplyr::filter(visit_code == "1500", .feature == "crispatus_dominance_by_mg")
  ) |> 
  mutate(crisp_dom_by_5week = outcome |> replace_na(FALSE)) 
  

```



**Visualization of the primary outcome** (and a few selected secondary outcomes)

```{r}
#| fig-width: 10
#| fig-height: 1.8

col_by_week5 |> 
  arrange(site, arm, LBP_colonization_by_week5) |>
  group_by(site, arm) |> 
  mutate(participant_number = row_number() |> factor()) |> 
  ungroup() |> 
  ggplot(aes(x = participant_number, y = site |> fct_rev(), col = site, fill = site, shape = LBP_colonization_by_week5)) +
  geom_point(size = 3) +
  facet_grid(. ~ arm, scales = "free_x", space = "free_x") +
  guides(color = "none", fill = "none") +
  scale_shape_manual("Colonized by LBP by week 5", values = c(1, 16)) +
  scale_color_manual(values = site_colors) + 
  xlab("Participant number") + ylab("") +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)
  ) 


```

```{r}
#| fig-width: 10
#| fig-height: 2


col_by_week5_qPCR |> 
  arrange(site, arm, LBP_colonization_by_week5) |>
  group_by(site, arm) |> 
  mutate(participant_number = row_number() |> factor()) |> 
  ungroup() |> 
  ggplot(aes(x = participant_number, y = site |> fct_rev(), col = site, fill = site, shape = LBP_colonization_by_week5)) +
  geom_point(size = 3) +
  facet_grid(. ~ arm, scales = "free_x", space = "free_x") +
  guides(color = "none", fill = "none") +
  scale_shape_manual("Colonized by LBP by week 5", values = c(1, 16)) +
  scale_color_manual(values = site_colors) + 
  xlab("Participant number") + ylab("") +
  ggtitle("Colonization based on the relative abundances estimated by qPCR") +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), 
    plot.title = element_text(hjust = 0.5)
  ) 

```


```{r}
#| fig-width: 10
#| fig-height: 2


detected_by_week5_qPCR |> 
  arrange(site, arm, LBP_detected_by_week5) |>
  group_by(site, arm) |> 
  mutate(participant_number = row_number() |> factor()) |> 
  ungroup() |> 
  ggplot(aes(x = participant_number, y = site |> fct_rev(), col = site, fill = site, shape = LBP_detected_by_week5)) +
  geom_point(size = 3) +
  facet_grid(. ~ arm, scales = "free_x", space = "free_x") +
  guides(color = "none", fill = "none") +
  scale_shape_manual("Colonized by LBP by week 5", values = c(1, 16)) +
  scale_color_manual(values = site_colors) + 
  xlab("Participant number") + ylab("") +
  ggtitle("At least 2 LBP strains detected at the same visit by week 5 qPCR") +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), 
    plot.title = element_text(hjust = 0.5)
  ) 

```


```{r}
#| fig-width: 10
#| fig-height: 2

crisp_dom_by_5week |> 
  arrange(site, arm, crisp_dom_by_5week) |>
  group_by(site, arm) |> 
  mutate(participant_number = row_number() |> factor()) |> 
  ungroup() |> 
  ggplot(aes(x = participant_number, y = site |> fct_rev(), col = site, fill = site, shape = crisp_dom_by_5week)) +
  geom_point(size = 3) +
  facet_grid(. ~ arm, scales = "free_x", space = "free_x") +
  guides(color = "none", fill = "none") +
  scale_shape_manual("Colonized by LBP by week 5", values = c(1, 16)) +
  scale_color_manual(values = site_colors) + 
  xlab("Participant number") + ylab("") +
  ggtitle("L. crispatus dominance (≥ 50% by metagenomics)") +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5), 
    plot.title = element_text(hjust = 0.5)
  ) 


```


```{r}

summary_table <- 
  col_by_week5 |> 
  group_by(site, arm) |> 
  summarize(
    n = n(), 
    n_success = sum(LBP_colonization_by_week5),
    .groups = "drop"
  ) |> 
  mutate(
    p = n_success / n,
    CI = binom::binom.confint(n_success, n, method = "exact")
  )

```

```{r}

table_2 <- 
  summary_table |> 
  mutate(
    N_per_site_and_arm = str_c("N = ", n),
    LBP_strain_detected = str_c(n_success, " (", round(p * 100)," %)"),
    CI_text = str_c(round(CI$lower * 100), "% - ", round(CI$upper * 100), "%")
  ) |> 
  dplyr::select(
    site, arm, N_per_site_and_arm, LBP_strain_detected, CI_text
  ) |> 
  pivot_longer(-c(site, arm)) |>
  mutate(
    name = 
      case_when(
        name == "N_per_site_and_arm" ~ "N participants",
        name == "LBP_strain_detected" ~ "n (%) participants\nwith LBP strain detected",
        name == "CI_text" ~ "95% CI"
      )
  ) |> 
  pivot_wider(
    id_cols = c(site, name), 
    names_from = arm,
    values_from = value, values_fill = ""
    )

```



```{r}

table_2 |> 
  group_by(site) |> 
  gt(caption = "Table 1: Colonization by week 5 by arm and site", row_group_as_column = TRUE) |> 
  cols_width(
    "name" ~ px(200),
    everything() ~ px(120)
  ) |> 
  cols_label(
    name = "",
    # Blinded = "All blinded arms",
    Pl = "Placebo",
    `LC-106-7` = "LC-106<br>7 days",
    `LC-106-3` = "LC-106<br>3 days",
    `LC-106-o` = "LC-106<br>overlap",
    `LC-115` = "LC-115<br>7 days",
    .fn = md
  )
  
```


#### Tests 

We test for differences between the placebo arm and each active arm using Fisher's exact test (exhaustive permutation test). 

Data from both site are merged for each arms given the low numbers of participants at MGH.

The $p$-values are adjusted for multiple testing using the Benjamini-Hochberg method.


```{r}

arms_to_test <- col_by_week5$arm |> unique() |> setdiff("Pl")

fisher_test_results <-
  map(
  arms_to_test,
  ~ {
    arm_to_test <- .x
    tmp <- col_by_week5 |> filter(arm %in% c("Pl", arm_to_test))
    table_for_test <- table(tmp$LBP_colonization_by_week5, tmp$arm |> fct_drop())
    test_res <- fisher.test(table_for_test, alternative = "greater")
    tibble(
      arm = arm_to_test,
      p_value = test_res$p.value,
      OR = test_res$estimate,
      lower_CI = test_res$conf.int[1],
      upper_CI = test_res$conf.int[2]
    )
  }
) |> 
  bind_rows() |> 
  mutate(
    adj_p_value = p_value |> p.adjust(method = "BH")
  )
  
```


```{r}

test_results <- 
  fisher_test_results |> 
  mutate(
    adj_p_value = adj_p_value |> scales::pvalue(accuracy = 0.001),
    OR = OR |> round(2) |> as.character(),
    CI = str_c(lower_CI |> round(2), " - ", upper_CI |> round(2))
    ) |> 
  select(arm, adj_p_value, OR, CI) |>
  pivot_longer(cols = -arm, names_to = "name", values_to = "value") |>
  pivot_wider(names_from = arm, values_from = value) |> 
  mutate(
    name = 
      case_when(
        name == "adj_p_value" ~ "Adj. p-value",
        name == "OR" ~ "OR",
        name == "CI" ~ "95% CI",
        TRUE ~ name
      )
  )
  
```

```{r}

table_2_with_model_results <- 
  bind_rows(
    table_2,
    test_results
  ) |> 
  mutate(
    site = site |> str_replace_na(""),
    Pl = 
      case_when(
        name == "Adj. p-value" ~ "Ref.",
        name == "OR" ~ "",
        str_detect(name, "CI") ~ "",
        TRUE ~ Pl
      )
  )

```


```{r}

table_2_with_model_results |> 
  group_by(site) |>
  gt(caption = "Table 2: Colonization by week 5 by arm and site", row_group_as_column = TRUE) |> 
  cols_width(
    "name" ~ px(150),
    everything() ~ px(120)
  ) |> 
  cols_label(
    name = "",
    Pl = "Placebo",
    `LC-106-7` = "LC-106<br>7 days",
    `LC-106-3` = "LC-106<br>3 days",
    `LC-106-o` = "LC-106<br>overlap",
    `LC-115` = "LC-115<br>7 days",
    .fn = md
  )
  
```





#### Model


```{r}

col_by_week5_agg <- 
  col_by_week5 |>
  group_by(site, arm) |>
  summarise(success = LBP_colonization_by_week5 |> sum(), ntrials = n()) 

bfit <- 
  brm(
    success | trials(ntrials) ~ arm + site + site:arm, 
    data = col_by_week5_agg, 
    family = binomial("logit")
    )

```



```{r}
#| eval: false
summary(bfit)
plot(bfit, ask = FALSE)
bind_cols(col_by_week5_agg, predict(bfit))
# loo(bfit, moment_match = TRUE, reloo = TRUE)
# pp_check(bfit)
```

```{r}
#| output: false

plot_res <- plot(conditional_effects(bfit), ask = FALSE)

```



```{r}
#| fig-width: 10
#| fig-height: 4

plot_res <- lapply(plot_res, function(p) {
  p + scale_color_manual(values = site_colors)
})

Reduce(`+`, plot_res) + 
  plot_layout(widths = c(2.2, 1, 3))

```




```{r}
#| eval: false

model_results <- summary(bfit)$fixed
model_results <- 
  model_results |> 
  rownames_to_column("term") |> 
  as_tibble() |> 
  mutate(
    factor = 
      case_when(
        str_detect(term, ":site") ~ "Arm x Site",
        str_detect(term, "^site") ~ "Site",
        str_detect(term, "^arm") ~ "Arm",
        TRUE ~ "Ref.",
      ),
    factor_level = 
      term |> 
      str_remove("arm") |> str_remove("site") |> str_replace(":"," x ") |> 
      str_replace("6M", "6-") |> str_replace("CM", "C-")
  ) |> 
  mutate(
    OR = ifelse(factor == "Ref.", NA, exp(Estimate)),
    lower_CI = ifelse(factor == "Ref.", NA, exp(`l-95% CI`)),
    upper_CI = ifelse(factor == "Ref.", NA, exp(`u-95% CI`))
  ) 

```

```{r}
#| eval: false

model_results |> 
  select(factor, factor_level, OR, lower_CI, upper_CI) |>
  gt() |> 
  gt::fmt_number(
    decimals = 2, n_sigfig = 3
  )

```


```{r}
#| eval: false

model_results_arm <- 
  model_results |> 
  filter(factor == "Arm") |> 
  mutate(
    OR = OR |> round(2) |> as.character(),
    CI = str_c(lower_CI |> round(2), " - ", upper_CI |> round(2))
    ) |> 
  select(factor_level, OR, CI) |>
  pivot_longer(cols = -factor_level, names_to = "name", values_to = "value") |>
  pivot_wider(names_from = factor_level, values_from = value) 
  
```

```{r}
#| eval: false

table_2_with_model_results <- 
  bind_rows(
    table_2,
    model_results_arm
  ) |> 
  mutate(
    site = site |> str_replace_na(""),
    Pl = 
      case_when(
        name == "OR" ~ "Ref.",
        name == "CI" ~ "",
        TRUE ~ Pl
      )
  )

```




```{r}
#| eval: false

table_2_with_model_results |> 
  group_by(site) |>
  gt(caption = "Table 2: Colonization by week 5 by arm and site", row_group_as_column = TRUE) |> 
  cols_width(
    "name" ~ px(200),
    everything() ~ px(120)
  ) |> 
  cols_label(
    name = "",
    Pl = "Placebo",
    `LC-106-7` = "LC-106<br>7 days",
    `LC-106-3` = "LC-106<br>3 days",
    `LC-106-o` = "LC-106<br>overlap",
    `LC-115` = "LC-115<br>7 days",
    .fn = md
  )
  
```


