---
title: "Processing Raw Data"
author: Laura Symul, Laura Vermeren, Susan Holmes
format: 
   html:
     page-layout: full
     code-fold: true
     toc: true
     toc-location: left
     toc-depth: 5
     embed-resources: true
execute:
  cache: true
  warning: false
knitr:
  opts_chunk:
    out.width: "100%"
editor: source
---

```{r}
#| warning: false

library(tidyverse)
library(magrittr)
library(gt)
library(patchwork)
library(SummarizedExperiment)
library(tidySummarizedExperiment)
library(brms)
library(gtsummary)
library(labelled)
library(scales)
# library(gridExtra)
library(ggpubr)
# library(mia) # BiocManager::install("mia")

theme_set(theme_light())
tmp <- fs::dir_map("R", source)

```

```{r}

#data_source <- "simulated"
data_source <- "real"

```

## Metagenomics 

### Importing Data 


```{r}

mg_readcount <- read_csv(str_c(data_dir(data_source), "MVIBR_kSanityVIRGO2_ReadCounts_20250404.csv"))
mg_relab <- read_csv(str_c(data_dir(data_source), "MVIBR_kSanityVIRGO2_relAbund_20250404.csv"))
mg_taxaGLcor <- read_csv(str_c(data_dir(data_source), "MVIBR_kSanityVIRGO2_taxaGLcor_20250404.csv"))
mg_technical_metadata <- read_csv(str_c(data_dir(data_source), "VIBRANT_MG_technicalMetaData_20250404.csv"))

```


Strain info 

```{r}

# strain info 
LBP_strain_info <- readxl::read_xlsx(str_c(dropbox_dir(), "9_VIBRANT_Study Product/IsolateNumbers.xlsx"))

LBP_strain_info <- 
 LBP_strain_info |> 
  mutate(strain_id = `VMRC ID`, 
         LBP = ifelse(is.na(LC106), "LC-115", "LC-106 & LC-115") |> factor(), 
         strain_origin = `Geographic source` |> factor()
  ) |> 
  dplyr::rename(Biose_ID = `Biose ID`, VMRC_ID = `VMRC ID`) |> 
  arrange(strain_origin, LBP) |> 
  mutate(strain_id = strain_id |> fct_inorder()) |> 
  dplyr::select(strain_id, LBP, strain_origin, contains("ID"))

LBP_strain_info |>   # gt(caption = "LBP strain information")
  set_variable_labels(
    strain_id = "Strain ID", 
    strain_origin = "Strain origin", 
    Biose_ID = "Biose ID", 
    VMRC_ID = "VMRC ID"
  ) |> 
  gt(caption = "LBP strain information") |> 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels())

```

The names of the strains do not correspond exactly to those in the metagenomics files. We will modify the names of the strains in *LBP_strain_info* to match them.

```{r}

LBP_strain_info <- 
  LBP_strain_info |>
  mutate(strain_id = sub("_.*", "", strain_id),
         strain_id = case_when(strain_id == "CC0028A1" ~ "C0028A1", 
                               TRUE ~ strain_id
    ))

LBP_strain_info |>   # gt(caption = "LBP strain information")
  set_variable_labels(
    strain_id = "Strain ID", 
    strain_origin = "Strain origin", 
    Biose_ID = "Biose ID", 
    VMRC_ID = "VMRC ID"
  ) |> 
  gt(caption = "LBP strain information") |> 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels())


```

Check that the column names (and therefore the species/strains) are the same in all 3 files.

```{r}

list(mg_readcount, mg_taxaGLcor, mg_relab) %>%
  purrr::map(~ colnames(.)) %>%
  purrr::reduce(~ setdiff(.x, .y)) %>%
  {if (length(.) > 0) cat("Colonnes différentes : ", ., "\n") else cat("Les noms de colonnes sont identiques.\n")}

# ou 

compare_columns <- function(df1, df2, df3) {
  cols1 <- colnames(df1)
  cols2 <- colnames(df2)
  cols3 <- colnames(df3)
  
  identical_cols <- all(cols1 %in% cols2) && all(cols2 %in% cols3) && all(cols1 %in% cols3)
  
  if (identical_cols) {
    return("The columns are identical in all three data frames.")
  } else {
    diff_cols <- list(
      data1 = setdiff(cols1, union(cols2, cols3)),
      data2 = setdiff(cols2, union(cols1, cols3)),
      data3 = setdiff(cols3, union(cols1, cols2))
    )
    return(diff_cols)
  }
}

result <- compare_columns(mg_readcount, mg_taxaGLcor, mg_relab)
print(result)


```


## Make Sumarrize Experiment object

>TODO : There's an extra column in `mg_readcount`. How do I manage this? Add an NA column?

```{r}

mg_to_SE <- function(mg_readcount,mg_taxaGLcor, mg_relab, mg_technical_metadata, LBP_strain_info){

  for (data in list(mg_readcount, mg_taxaGLcor, mg_relab, mg_technical_metadata)) {
    if (any(duplicated(data$sampleID))) stop("Duplicated `uid` in `mg`")
  }
  
  # create assays
  assay_readcounts <- 
    mg_readcount |> 
    dplyr::select(-c(CST, subCST, score)) |>
    as.data.frame() |> 
    column_to_rownames("sampleID") |>
    t() 

  assay_readcount_corr <- 
    mg_taxaGLcor |> 
    dplyr::select(-c(CST, subCST, score)) |>
    as.data.frame() |>
    column_to_rownames("sampleID") |>
    t()
  
  assay_relative_ab <- 
    mg_relab |> 
    dplyr::select(-c(CST, subCST, score)) |>
    as.data.frame() |> 
    column_to_rownames("sampleID") |>
    t()

  se_coldata <-
    mg_technical_metadata |>
    rename(sampleID = UID) |>
    as.data.frame()

  se_rowdata <- 
    mg_readcount |> 
    dplyr::select(-c(sampleID, CST, subCST, score)) |>
    colnames() |> 
    as.data.frame() |> 
    setNames("strain") |> 
    distinct() |> 
    left_join(
      LBP_strain_info,
      by = join_by(strain == strain_id)
    ) 
  
  # Harmonization of the order of samples and feature
  
  sorted_sample_ids <- sort(as.character(se_coldata$sampleID))
  assay_readcounts <- assay_readcounts[, sorted_sample_ids]
  assay_readcount_corr <- assay_readcount_corr[, sorted_sample_ids]
  assay_relative_ab <- assay_relative_ab[, sorted_sample_ids]
  se_coldata <- se_coldata[match(sorted_sample_ids, se_coldata$sampleID), ]

  sorted_strains <- assay_relative_ab |> rownames() # sort(as.character(se_rowdata$strain))
  assay_readcounts <- assay_readcounts[sorted_strains, ]
  assay_readcount_corr <- assay_readcount_corr[sorted_strains, ]
  assay_relative_ab <- assay_relative_ab[sorted_strains, ]
  se_rowdata <- se_rowdata[match(sorted_strains, se_rowdata$strain), ]

  SummarizedExperiment::SummarizedExperiment(
    assays = list(readcounts = assay_readcounts, readcount_corr = assay_readcount_corr, relative_ab = assay_relative_ab),
    rowData = se_rowdata,
    colData = se_coldata
  )
}

```

```{r}

# SE_mg <- mg_to_SE(mg_readcount, mg_taxaGLcor, mg_relab, mg_technical_metadata, LBP_strain_info)

```


### Checks & Quality Controls

#### Link between readcount relatie abundance 

We test whether the relative abundance is equal to the readcount divided by the total number of reads. To do this we test :
for the uncorrected readcount and the corrected red count if :

- relative abundance (all strains/species) = readcount / total reads
- relative abundance (LBP strains) = readcount / total reads LBP
- relative abundance (LBP starin) = readcount / total reads LBP + crispatus


```{r}

mg_calc_relab <- 
  mg_readcount |> 
  mutate(
    sum_read_LBP = rowSums(across(all_of(LBP_strain_info$strain_id))),
    sum_read_LBPandCripsatus = rowSums(across(c(all_of(LBP_strain_info$strain_id), Lactobacillus_crispatus))),
    sum_read_total = rowSums(across(-c(sampleID, CST, subCST, score)))
  ) |> 
  select(-c(CST, subCST, score)) |>
  pivot_longer(
    -c(sampleID, sum_read_LBP, sum_read_LBPandCripsatus, sum_read_total), 
    names_to = "taxon", 
    values_to = "readcount") |>
  mutate(
    calc.rel_ab_total = readcount / sum_read_total,
    calc.rel_ab_LBP = readcount / sum_read_LBP,
    calc.rel_ab_LBPandCripsatus = readcount / sum_read_LBPandCripsatus, 
    calc.rel_ab_without_LBP = readcount / (sum_read_total - sum_read_LBP),
    calc.rel_ab_without_LBPandCrispatus = readcount / (sum_read_total - sum_read_LBPandCripsatus)
  ) 

mg_calc_relab <-
  mg_calc_relab |>
  left_join(
    mg_taxaGLcor |>
      mutate(
        sum_read.corr_LBP = rowSums(across(all_of(
          LBP_strain_info$strain_id
        ))),
        sum_read.corr_LBPandCripsatus = rowSums(across(c(
          all_of(LBP_strain_info$strain_id), Lactobacillus_crispatus
        ))),
        sum_read.corr_total = rowSums(across(-c(
          sampleID, CST, subCST, score
        )))
      ) |>
      select(-c(CST, subCST, score)) |>
      pivot_longer(
        -c(
          sampleID,
          sum_read.corr_LBP,
          sum_read.corr_LBPandCripsatus,
          sum_read.corr_total
        ),
        names_to = "taxon",
        values_to = "readcount_corr"
      ) |>
      mutate(
        calc.rel_ab_total.corr = readcount_corr / sum_read.corr_total,
        calc.rel_ab_LBP.corr = readcount_corr / sum_read.corr_LBP,
        calc.rel_ab_LBPandCripsatus.corr = readcount_corr / sum_read.corr_LBPandCripsatus,
        calc.rel_ab_without_LBP.corr = readcount_corr / (sum_read.corr_total - sum_read.corr_LBP),
        calc.rel_ab_without_LBPandCrispatus.corr = readcount_corr / (sum_read.corr_total - sum_read.corr_LBPandCripsatus)
      ),
    by = c("sampleID", "taxon")
  )


mg_calc_relab <- 
  mg_calc_relab |> 
  left_join(
    mg_relab |> 
      select(-c(CST, subCST, score)) |> 
      pivot_longer(
        -c(sampleID), 
        names_to = "taxon", 
        values_to = "rel_ab"
      ),
    by = c("sampleID", "taxon")
  ) 

# rel_ab = readcount / sum_read_total
mg_calc_relab |> 
  ggplot(aes(x= rel_ab, y = calc.rel_ab_total)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") + 
  labs(x = "Relative abundance", y = "Calculated relative abundance \n (read count/ total read count)", 
       title = "Uncorrected readcount")


# mg_calc_relab |> 
#   ggplot(aes(x= rel_ab, y = calc.rel_ab_total, col = sampleID)) +
#   geom_point(alpha=0.5) +
#   geom_abline(slope = 1, intercept = 0, col = "red") + 
#   labs(x = "Relative abundance", y = "Calculated relative abundance \n (read count/ total read count)", 
#        title = "Uncorrected readcount") 


mg_calc_relab |>
  filter(taxon %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x= rel_ab, y = calc.rel_ab_LBP)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Relative abundance for LBP strain", y = "Calculated relative abundance \n (read count/ LBP read count)", 
       title = "Uncorrected readcount")

mg_calc_relab |>
  filter(taxon %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x= rel_ab, y = calc.rel_ab_LBPandCripsatus)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Relative abundance for LBP strain", y = "Calculated relative abundance \n (read count/ LBP & crispatus read count)", title = "Uncorrected readcount")

mg_calc_relab |> 
  filter(!taxon %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x= rel_ab, y = calc.rel_ab_without_LBP)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Relative abundance", 
       y = "Calculated relative abundance (species without LBP) \n (read count/ total read count (without LBP)", 
       title = "Uncorrected readcount")

mg_calc_relab |> 
  filter(!taxon %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x= rel_ab, y = calc.rel_ab_without_LBPandCrispatus)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Relative abundance", 
       y = "Calculated relative abundance (species without LBP) \n (read count/ total read count (without LBP & crisp.))", 
       title = "Uncorrected readcount")


mg_calc_relab |> 
  ggplot(aes(x= rel_ab, y = calc.rel_ab_total.corr)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") + 
  labs(x = "Relative abundance", y = "Calculated relative abundance \n (read count/ total read count)", 
       title = "Corrected readcount")

mg_calc_relab |>
  filter(taxon %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x= rel_ab, y = calc.rel_ab_LBP.corr)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Relative abundance for LBP strain", y = "Calculated relative abundance \n (read count/ LBP read count)", 
       title = "Corrected readcount")

mg_calc_relab |>
  filter(taxon %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x= rel_ab, y = calc.rel_ab_LBPandCripsatus.corr)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Relative abundance for LBP strain", y = "Calculated relative abundance \n (read count/ LBP & crispatus read count)", title = "Corrected readcount")

mg_calc_relab |> 
  filter(!taxon %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x= rel_ab, y = calc.rel_ab_without_LBP.corr)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Relative abundance", 
       y = "Calculated relative abundance (species without LBP) \n (read count/ total read count (without LBP)", 
       title = "Corrected readcount")

mg_calc_relab |> 
  filter(!taxon %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x= rel_ab, y = calc.rel_ab_without_LBPandCrispatus.corr)) +
  geom_point(alpha=0.5) +
  geom_abline(slope = 1, intercept = 0, col = "red") +
  labs(x = "Relative abundance", 
       y = "Calculated relative abundance (species without LBP) \n (read count/ total read count (without LBP & crisp.))", 
       title = "Corrected readcount")
  

  

```

**Total readcounts per sample**

```{r}

mg_readcount_long <- 
  mg_readcount |> 
  select(-c(CST, subCST, score)) |>
  pivot_longer(-c(sampleID), names_to = "taxon", values_to = "readcounts")

```


```{r}

mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  ggplot(aes(x = total_count)) +
  geom_histogram() +
  scale_x_log10()

```


```{r}

total_reads <- 
 mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  left_join(
    mg_technical_metadata |> dplyr::rename(sampleID = UID) 
  )


total_reads |> 
  ggplot(aes(x = total_count, fill = SampleType)) +
  geom_histogram() +
  scale_x_log10()

```
Proportion of 0
```{r}
# proportion of zero 
(mg_readcount_long$readcounts == 0)  |> mean()
```

```{r}

mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  ggplot(aes(x = total_count, y = sampleID |> fct_rev())) +
  geom_col() +
  ylab("Sample ID") + xlab("Total number of counts") 

```


**Total corrected readcounts per sample**

```{r}

mg_readcount.corr_long <- 
  mg_taxaGLcor |> 
  select(-c(CST, subCST, score)) |>
  pivot_longer(-c(sampleID), names_to = "taxon", values_to = "readcounts")

```


```{r}

mg_readcount.corr_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  ggplot(aes(x = total_count)) +
  geom_histogram() +
  scale_x_log10()

```

Proportion of 0

```{r}

(mg_taxaGLcor_long$readcounts_corr == 0)  |> mean()

```


**Total proportion per sample**

```{r}

mg_relab_long <- 
  mg_relab |> 
  select(-c(CST, subCST, score)) |>
  pivot_longer(-c(sampleID), names_to = "taxon", values_to = "relab")

```


```{r}

mg_relab_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(relab)) |> 
  ggplot(aes(x = total_count)) +
  geom_histogram() +
  scale_x_log10()

```




***Missing Values**

```{r}

mg_readcount_long |>
  ggplot(aes(x=taxon, y = sampleID, fill = is.na(readcounts))) + 
  geom_tile() + 
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  labs(title = "Missing values in readcount") +
  scale_fill_manual(
    values = c("TRUE" = "pink", "FALSE" = "steelblue2"),  
    labels = c("FALSE" = "Présent", "TRUE" = "Manquant")) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mg_readcount.corr_long |>
  ggplot(aes(x=taxon, y = sampleID, fill = is.na(readcounts_corr))) + 
  geom_tile() + 
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  labs(title = "Missing values in corrected readcount") +
  scale_fill_manual(
    values = c("TRUE" = "pink", "FALSE" = "steelblue2"),  
    labels = c("FALSE" = "Présent", "TRUE" = "Manquant")) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mg_relab_long |>
  ggplot(aes(x=taxon, y = sampleID, fill = is.na(relab))) + 
  geom_tile() + 
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  labs(title = "Missing values in relative abundance") +
  scale_fill_manual(
    values = c("TRUE" = "pink", "FALSE" = "steelblue2"),  
    labels = c("FALSE" = "Présent", "TRUE" = "Manquant")) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))



```
**Outliers**

```{r}

one_third <- floor((ncol(mg_readcount) - 4)/3)
two_third <- floor((ncol(mg_readcount) - 4)/3*2)
three_third <- ncol(mg_readcount) - 4

one_quarter <- floor((ncol(mg_readcount) - 4)/4)
two_quarter <- floor((ncol(mg_readcount) - 4)/4*2)
three_quarter <- floor((ncol(mg_readcount) - 4)/4*3)
fourth_quarter <- ncol(mg_readcount) - 4


mg_readcount |> 
  select(-c(CST, subCST, score)) |>
  select(1, 2:one_quarter) |> 
  pivot_longer(-sampleID, names_to = "taxon", values_to = "readcounts") |>
  ggplot(aes(x = taxon, y = readcounts)) + 
  geom_boxplot() + 
  scale_x_discrete("Strains") +
  scale_y_log10() +
  labs(title = "Outliers in readcount - Partie 1") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mg_readcount |> 
  select(-c(CST, subCST, score)) |>
  select(1, one_quarter:two_quarter) |> 
  pivot_longer(-sampleID, names_to = "taxon", values_to = "readcounts") |>
  ggplot(aes(x = taxon, y = readcounts)) + 
  geom_boxplot() + 
  scale_x_discrete("Strains") +
  scale_y_log10() +
  labs(title = "Outliers in readcount - Partie 1") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mg_readcount |> 
  select(-c(CST, subCST, score)) |>
  select(1, two_quarter:three_quarter) |> 
  pivot_longer(-sampleID, names_to = "taxon", values_to = "readcounts") |>
  ggplot(aes(x = taxon, y = readcounts)) + 
  geom_boxplot() + 
  scale_x_discrete("Strains") +
  scale_y_log10() +
  labs(title = "Outliers in readcount - Partie 1") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

mg_readcount |> 
  select(-c(CST, subCST, score)) |>
  select(1, three_quarter:fourth_quarter) |> 
  pivot_longer(-sampleID, names_to = "taxon", values_to = "readcounts") |>
  ggplot(aes(x = taxon, y = readcounts)) + 
  geom_boxplot() + 
  scale_x_discrete("Strains") +
  scale_y_log10() +
  labs(title = "Outliers in readcount - Partie 1") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))


```


**Proportion of LBP per sample** 

```{r}
#| fig-height: 10

mg_relab_long |> 
  filter(taxon %in% all_of(LBP_strain_info$strain_id)) |> 
  left_join(LBP_strain_info, by = c("taxon" = "strain_id")) |>
  ggplot(aes(x = taxon, y = sampleID, fill = relab)) +
  geom_tile() +
  scale_fill_continuous(low = "white", high = "steelblue2") +
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  facet_grid(. ~ LBP + strain_origin, scales = "free_x", space = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r}
#| fig-height: 15

mg_relab_long |> 
  filter(taxon %in% all_of(LBP_strain_info$strain_id)) |> 
  left_join(LBP_strain_info, by = c("taxon" = "strain_id")) |>
  group_by(taxon) |> 
  mutate(max_rel_ab = max(relab), rel_rel_ab = relab/max_rel_ab) |>
  ungroup() |> 
  ggplot(aes(x = taxon, y = sampleID, fill = rel_rel_ab)) +
  geom_tile() +
  scale_fill_continuous(low = "white", high = "steelblue2") +
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  facet_grid(. ~ LBP + strain_origin, scales = "free_x", space = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


```{r}
#| fig-height: 15

mg_relab_long |> 
  filter(taxon %in% all_of(LBP_strain_info$strain_id)) |> 
  left_join(LBP_strain_info, by = c("taxon" = "strain_id")) |>
  group_by(taxon) |> 
  mutate(
    max_rel_ab = max(relab), 
    rel_rel_ab = relab/max_rel_ab,
    tot_rel_ab = sum(relab),
    f_rel_ab = sum(relab != 0)/n()
    ) |>
  ungroup() |> 
  arrange(f_rel_ab) |> 
  mutate(taxon = taxon |> fct_inorder()) |> 
  arrange(taxon) |> 
  mutate(sampleID = sampleID |> fct_inorder()) |> 
  ggplot(aes(x = taxon, y = sampleID, fill = (relab == 0))) +
  geom_tile() +
  # scale_fill_continuous(low = "white", high = "steelblue2") +
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  facet_grid(. ~ LBP + strain_origin, scales = "free_x", space = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```





**Proportion of all species by sample**

```{r}

mg_relab_long |> 
  filter(!taxon %in% LBP_strain_info$strain_id) |> 
  left_join(LBP_strain_info, by = c("taxon" = "strain_id")) |>
  ggplot(aes(x = taxon, y = sampleID, fill = relab)) +
  geom_tile() +
  scale_fill_continuous(low = "white", high = "steelblue2") +
  scale_x_discrete("Species (without LBP") +
  scale_y_discrete("Samples",breaks = NULL) +
  facet_grid(. ~ LBP + strain_origin, scales = "free_x", space = "free") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

#### check by technical data


```{r}

mg_relab_long <- 
  mg_relab |> 
  left_join(
    mg_technical_metadata |>
      filter(is.na(LibrarySequencedTwice)) |> 
      select(UID, Ext_Lib_Plate, Ext_Lib_Position, Lane, FragmentSize, BioinformaticsProcessingBatch),
    by = c("sampleID" = "UID")
  ) |> 
  pivot_longer(
    -c(sampleID, CST, subCST, score, Ext_Lib_Plate, Ext_Lib_Position, Lane, FragmentSize, BioinformaticsProcessingBatch), 
    names_to = "strain_id", 
    values_to = "rel_ab"
  )

```

**Relative abundance of LBP strain by plate**

```{r}

mg_relab_long |> 
  filter(strain_id %in% LBP_strain_info$strain_id) |>
  ggplot(aes(x = strain_id, y = sampleID, fill = rel_ab)) +
  geom_tile() +
  scale_fill_continuous(low = "white", high = "steelblue2") +
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  facet_wrap(~Ext_Lib_Plate, scales = "free_y", strip.position = "top") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

**Relative abundance of LBP strain by batch**

```{r}

mg_relab_long |> 
  filter(strain_id %in% LBP_strain_info$strain_id) |> 
  ggplot(aes(x = strain_id, y = sampleID, fill = rel_ab)) +
  geom_tile() +
  scale_fill_continuous(low = "white", high = "steelblue2") +
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  facet_wrap(~ BioinformaticsProcessingBatch, scales = "free_y", strip.position = "top") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

**Relative abundance of LBP strain by lane**

```{r}

mg_relab_long |> 
  filter(strain_id %in% LBP_strain_info$strain_id) |> 
  ggplot(aes(x = strain_id, y = sampleID, fill = rel_ab)) +
  geom_tile() +
  scale_fill_continuous(low = "white", high = "steelblue2") +
  scale_x_discrete("Strains") +
  scale_y_discrete("Samples",breaks = NULL) +
  facet_wrap(~ Lane, scales = "free_y", strip.position = "top") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


**Total number of counts per sample by plate**


```{r}

mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  left_join(
    mg_technical_metadata |>
      filter(is.na(LibrarySequencedTwice)) |> 
      select(UID, Ext_Lib_Plate, Ext_Lib_Position, Lane, FragmentSize, BioinformaticsProcessingBatch),
    by = c("sampleID" = "UID")
  )  |>  
  ggplot(aes(x = total_count)) +
  geom_histogram() +
  ylab("Sample ID") + xlab("Total number of counts") + 
  facet_grid(Ext_Lib_Plate ~ ., scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_log10()

```

```{r}

mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  left_join(
    mg_technical_metadata |>
      filter(is.na(LibrarySequencedTwice)) |> 
      select(UID, Ext_Lib_Plate, Ext_Lib_Position, Lane, FragmentSize, BioinformaticsProcessingBatch),
    by = c("sampleID" = "UID")
  )  |>  
  ggplot(aes(x = total_count, y = sampleID |> fct_rev())) +
  geom_col() +
  ylab("Sample ID") + xlab("Total number of counts") + 
  facet_wrap(~ Ext_Lib_Plate, scales = "free_y", strip.position = "top") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


**Total number of counts per sample by batch**

```{r}

mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  left_join(
    mg_technical_metadata |>
      filter(is.na(LibrarySequencedTwice)) |> 
      select(UID, Ext_Lib_Plate, Ext_Lib_Position, Lane, FragmentSize, BioinformaticsProcessingBatch),
    by = c("sampleID" = "UID")
  )  |>  
  ggplot(aes(x = total_count)) +
  geom_histogram() +
  ylab("Sample ID") + xlab("Total number of counts") + 
  facet_grid(BioinformaticsProcessingBatch ~ ., scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_log10()

```



```{r}

mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  left_join(
    mg_technical_metadata |>
      filter(is.na(LibrarySequencedTwice)) |> 
      select(UID, Ext_Lib_Plate, Ext_Lib_Position, Lane, FragmentSize, BioinformaticsProcessingBatch),
    by = c("sampleID" = "UID")
  )  |>  
  ggplot(aes(x = total_count, y = sampleID |> fct_rev())) +
  geom_col() +
  ylab("Sample ID") + xlab("Total number of counts") + 
  facet_wrap(~ BioinformaticsProcessingBatch, scales = "free_y", strip.position = "top") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```
**Total number of counts per sample by lane**



```{r}

mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  left_join(
    mg_technical_metadata |>
      filter(is.na(LibrarySequencedTwice)) |> 
      select(UID, Ext_Lib_Plate, Ext_Lib_Position, Lane, FragmentSize, BioinformaticsProcessingBatch),
    by = c("sampleID" = "UID")
  )  |>  
  ggplot(aes(x = total_count)) +
  geom_histogram() +
  ylab("Sample ID") + xlab("Total number of counts") + 
  facet_grid(Lane ~ ., scales = "free_y") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_log10()

```



```{r}
mg_readcount_long |> 
  group_by(sampleID) |> 
  summarise(total_count = sum(readcounts)) |> 
  left_join(
    mg_technical_metadata |>
      filter(is.na(LibrarySequencedTwice)) |> 
      select(UID, Ext_Lib_Plate, Ext_Lib_Position, Lane, FragmentSize, BioinformaticsProcessingBatch),
    by = c("sampleID" = "UID")
  )  |>  
  ggplot(aes(x = total_count, y = sampleID |> fct_rev())) +
  geom_col() +
  ylab("Sample ID") + xlab("Total number of counts") + 
  facet_wrap(~ Lane, scales = "free_y", strip.position = "top") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

